Data analysis of dangerous dog attack

### 2024-09-25 paper/paper.qmd
draft paper

---
title: "Data analysis of dangerous dog attack"
subtitle: "Analyzing the circumstance of dog attack and compare data to find out the most dangerous breed"
author: 
  - Yuanchen Miao
  - Another author
thanks: "Code and data are available at: https://github.com/NevaeH-9/dangerous-dog-attack.git."
date: 24 september 2024
date-format: long
abstract: "Nowadays, almost every family choose to have a pet, mostly dogs. Dogs have their innate nature to protect themselves or their owner from the danger they meet, sometimes they bark at it, sometimes they make an attack on it. Dogs require training from human to prevent them from attacking or barking at people, but some breed of dogs have stronger aggressive desire, in some situations, these breed of dogs probably attack people after training. In this report, data of Registry of Dogs Subject to and Issued a Dangerous Dog Order from Open Data Toronto is used, this data is refreshed monthly to today and all died dogs are removed from the data. There appears some specific breed of dogs are more likely to attack people and causing worse attck circumstance."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(here)
```

@@

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

@@

>>>>

---
title: "Data analysis of dangerous dog attack"
subtitle: "Analyzing the circumstance of dog attack and compare data to find out the most dangerous breed"
author: 
  - Yuanchen Miao
thanks: "Code and data are available at: https://github.com/NevaeH-9/dangerous-dog-attack.git."
date: 24 september 2024
date-format: long
abstract: "Nowadays, almost every family choose to have a pet, mostly dogs. Dogs have their innate nature to protect themselves or their owner from the danger they meet, sometimes they bark at it, sometimes they make an attack on it. Dogs require training from human to prevent them from attacking or barking at people, but some breed of dogs have stronger aggressive desire, in some situations, these breed of dogs probably attack people after training. In this report, data of Registry of Dogs Subject to and Issued a Dangerous Dog Order from Open Data Toronto is used, this data is refreshed monthly to today and all died dogs are removed from the data. There appears some specific breed of dogs are more likely to attack people and causing worse attck circumstance."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(here)
library(dplyr)
```

@@

Since a lot of people are owning dogs nowadays and dogs have possibility to be triggered to attack people with a lot of different [reasons](https://www.calmcanineacademy.com/blog/2020/11/06/let-s-talk-about-dog-attacks), like barriers, resources, 'trigger stacking'[@calmcanine2020], there are a lot of people reporting they are getting attack by dogs every year. There are [2726 service requests](https://www.cbc.ca/news/canada/toronto/dangerous-dog-owners-new-rules-city-toronto-1.7183803#:~:text=Toronto%20Animal%20Services%20investigates%20all,increase%20from%202%2C653%20in%202022.) for potentially dog attack in '2023'[@cbc2023].

# Data {#sec-data}

## Overview {#sec-data-overview}
The dataset '"About registry of Dogs Subject to and Issued a Dangerous Dog Order" '[@toronto2023]has been used in this paper to analyze what breed of dogs are tend to be more aggressive and what breed of dogs is making worse attack circumstances. 'Programming language R'[@citeR] is been used in this paper to clean and analyze the dataset. Packages that could installed in R: 'tidyverse'[@tidyverse], 'ggplot2'[@ggplot], 'dplyr'[@dplyr], 'here'[@here] and ‘opendatatoronto'[@opendatatoronto]are used to download the dataset to R and making test to the dataset.

## Results {#sec-data-results}
```{r}
#| echo: false
#| warning: false
#| message: false

cleaned_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```
Loading data with the programming language 'R'[@citeR] and package 'here'[@here], graphs are generate by using the package 'tidyverse'[@tidyverse] and 'ggplot2'[@ggplot]. Codes are adapted from 'Tellingstorieswithdata' [@rohan]

@@

```{r}
#| label: fig-1
#| fig-cap: "Number of dogs that have been issued as a dangerous dog in Toronto" 
#| warning: false
#| message: false
#| echo: false

cleaned_data1 <- cleaned_data %>%
  mutate(Year = year(as.Date(Date_of_Dangerous_Act)))

cleaned_data1 |>
  ggplot(aes(x = Year)) +
  geom_bar() +
  theme_minimal() + 
  labs(title = "Number of dangerous dog order in Toronto by year",
       x = "Year", y = "Dangerous act"
       ) 
```
More dogs are issued as dangerous dog between year 2022 to 2024(@fig-1)

```{r}
#| label: fig-2
#| fig-cap: "Number of different attack circumstance " 
#| warning: false
#| message: false
#| echo: false

cleaned_data1 |>
  ggplot(aes(x = Bite_Circumstance)) +
  geom_bar() +
  theme_minimal() + 
  labs(title = "number of different bite circumstances",
       x = "Bite circumstances", y = "number of bite circumstances"
       ) 
```
The figure shows that the circumstances of bite by dogs (@fig-2)

```{r}
#| label: fig-3
#| fig-cap: "Breed of dogs that has dangerous act above the mean of dangerous dog order" 
#| warning: false
#| message: false
#| echo: false

breed_counts <- cleaned_data %>%
  group_by(Breed) %>%
  summarise(Num_Dangerous_Acts = n())

mean_value <- mean(breed_counts$Num_Dangerous_Acts)

filtered_breeds <- breed_counts %>%
  filter(Num_Dangerous_Acts > mean_value)

ggplot(data = filtered_breeds, aes(x = Breed, y = Num_Dangerous_Acts, fill = Breed)) +
  geom_bar(stat = "identity") +
  labs(title = "Dog Breeds with Dangerous Acts Above the Mean",
       x = "Breed of Dog",
       y = "Number of Dangerous Acts") +
  theme_minimal() +
  theme(
    legend.position = "none",axis.title.x = element_text(size = 1), axis.text.x = element_text(size = 6)
  )
```
If the breed of dog contains the gene of BULLDOG ro SHEPHERD, they are tend to be more aggressive.@fig-3

```{r}
#| label: fig-4
#| fig-cap: "Number of severe each breed of dog caused" 
#| warning: false
#| message: false
#| echo: false

severe_counts <- cleaned_data %>%
  filter(Bite_Circumstance == "SEVERE") %>%
  group_by(Breed) %>%
  summarise(Num_severe = n())

mean_value <- mean(severe_counts$Num_severe)

filtered_breeds1 <- severe_counts %>%
  filter(Num_severe > mean_value)

# Create the bar plot
ggplot(data = filtered_breeds1, aes(x = Breed, y = Num_severe, fill = Breed)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of 'severe' Reports by Dog Breed",
       x = "Breed of Dog",
       y = "Number of 'severe' Reports") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(size = 6))
```

```{r}
#| label: fig-5
#| fig-cap: "Number of very severe each breed of dog caused" 
#| warning: false
#| message: false
#| echo: false

very_severe <- cleaned_data %>%
  filter(Bite_Circumstance == "VERY SEVERE") %>%
  group_by(Breed) %>%
  summarise(Num_very = n())

# Create the bar plot
ggplot(data = very_severe, aes(x = Breed, y = Num_very, fill = Breed)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of 'VERY SEVERE' Reports by Dog Breed",
       x = "Breed of Dog",
       y = "Number of 'VERY SEVERE' Reports") +
  theme_minimal() +
  theme(legend.position = "none")
```
